### 2018-01-18 wfe/request/request-template.tex

% LaTeX template for Artifact Evaluation V20170622
%
% Prepared by Grigori Fursin (cTuning foundation, France and dividiti, UK),
% Bruce Childers (University of Pittsburgh, USA),
% Thierry Moreau (University of Washington, USA)
%
% See example of this Artifact Appendix in the CGO'17 paper (with the CK workflow framework):
% https://www.cl.cam.ac.uk/~sa614/papers/Software-Prefetching-CGO2017.pdf
%
% (C)opyright 2014-2017
%
% CC BY 4.0 license
%

\documentclass{sigplanconf}

\begin{document}

\special{papersize=8.5in,11in}

\title{Title}

\authorinfo{John Doe}
{Affiliation}
{johndoe@email.com}

\maketitle

\begin{abstract}
Summary of work, and artifact.
\end{abstract}

\section{Extended Abstract}

\subsection{Technical Description}
{\em Obligatory. Provide a technical desciption of your novel work, or references to
your already published work which you are providing an artifact for.}

\subsection{Empirical Evaluation}
{\em Obligatory. Provide empirical evaluation results of the technique together with improved metrics, 
optimization, or exploration which you are providing an artifact for, and discuss the insights 
that the evaluation provides. The methodology can be discussed in more details in the artifact appendix.}

\newpage

\appendix
\section{Artifact appendix}

Submission and reviewing guidelines and methodology: \\
{\em http://cTuning.org/ae/submission.html}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{Abstract}

{\em Obligatory}

\subsection{Artifact check-list (meta-information)}

{\em Obligatory. Use just a few informal keywords in all fields applicable to your artifacts
and remove the rest. This information is needed to find appropriate reviewers and gradually 
unify artifact meta information in Digital Libraries.}

{\small
\begin{itemize}
  \item {\bf Algorithm: }
  \item {\bf Program: }
  \item {\bf Compilation: }
  \item {\bf Transformations: }
  \item {\bf Binary: }
  \item {\bf Data set: }
  \item {\bf Run-time environment: }
  \item {\bf Hardware: }
  \item {\bf Run-time state: }
  \item {\bf Execution: }
  \item {\bf Metrics: }
  \item {\bf Output: }
  \item {\bf Experiments: }
  \item {\bf Workflow frameworks used?: }
  \item {\bf Publicly available?: }
\end{itemize}

{\em Obligatory. Provide brief assessment of your artifacts 
based on {\em https://www.acm.org/publications/policies/artifact-review-badging}}:

\begin{itemize}
  \item {\bf Artifacts publicly available?: }
  \item {\bf Artifacts functional?:}
  \item {\bf Artifacts reusable?:}
  \item {\bf Results validated?:}
\end{itemize}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{Description}

\subsubsection{How delivered}

{\em Obligatory}

\subsubsection{Hardware dependencies}

\subsubsection{Software dependencies}

\subsubsection{Data sets}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{Installation}

{\em Obligatory}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{Experiment workflow}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{Evaluation and expected result}

{\em Obligatory}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{Experiment customization}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{Notes}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% When adding this appendix to your paper, 
% please remove below part
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\end{document}