### 2018-02-02 wfe/request/request-template.tex
rewritten ReQuEST CFP +  updated template

% LaTeX template for Artifact Evaluation V20180201
%
% Prepared by Grigori Fursin (cTuning foundation, France and dividiti, UK)
% and Thierry Moreau (University of Washington, USA)
%
% This document was derived from the original Artifact Appendix
% prepared by Grigori Fursin (cTuning foundation, France and dividiti, UK) 
% and Bruce Childers (University of Pittsburgh, USA)
%
% See example of this Artifact Appendix in the CGO'17 paper (with the CK workflow framework):
% https://www.cl.cam.ac.uk/~sa614/papers/Software-Prefetching-CGO2017.pdf
%
% (C)opyright 2014-2018
%
% CC BY 4.0 license
%

\documentclass{sigplanconf}

\begin{document}

\special{papersize=8.5in,11in}

\title{Title}

\authorinfo{John Doe}
{Affiliation}
{johndoe@email.com}

\maketitle

\begin{abstract}
Summary of work and associated workflow artifact.

Please, specify if you want a public or a private evaluation.
\end{abstract}

\section{Extended Abstract}

\subsection{Technical Description}
{\em Obligatory. Provide a technical description of your novel work, 
or references to your already published work which you are providing an artifact for.}

\subsection{Empirical Evaluation}
{\em Obligatory. Provide empirical evaluation results of the technique together with improved metrics, 
optimization, or exploration which you are providing an artifact for, and discuss the insights 
that the evaluation provides. The methodology can be discussed in more details in the artifact appendix.}

\newpage

\appendix
\section{Artifact Appendix}

Submission and reviewing methodology: \\
{\em http://cTuning.org/ae/submission-20171101.html}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{Abstract}

{\em Obligatory}

\subsection{Artifact check-list (meta-information)}

{\em Obligatory. Use just a few informal keywords in all fields applicable to your artifacts
and remove the rest. This information is needed to find appropriate reviewers and gradually 
unify artifact meta information in Digital Libraries.}

{\small
\begin{itemize}
  \item {\bf Algorithm: }
  \item {\bf Program: }
  \item {\bf Compilation: }
  \item {\bf Transformations: }
  \item {\bf Binary: }
  \item {\bf Data set: }
  \item {\bf Run-time environment: }
  \item {\bf Hardware: }
  \item {\bf Run-time state: }
  \item {\bf Execution: }
  \item {\bf Metrics: }
  \item {\bf Output: }
  \item {\bf Experiments: }
  \item {\bf Collective Knowledge workflow framework used?: } 
  \item {\bf Publicly available?: }
\end{itemize}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{Description}

\subsubsection{How to obtain?}

{\em Obligatory}

\subsubsection{Hardware dependencies}

\subsubsection{Software dependencies}

\subsubsection{Data sets}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{Installation}

{\em Obligatory}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{Experiment workflow}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{Evaluation and expected result}

{\em Obligatory}

Please, specify which metrics you improve
(execution time, accuracy, power consumption, 
memory usage, device cost, usage cost, etc)

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{Experiment customization}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{Notes}

\end{document}